\section{The problem}
Dr. Giuseppe Maggiore has developed a language specifically tailored for game development, called Casanova. To create this language he also created a language called MetaCompiler (MC). MC is a language which focuses itself on building compilers for other languages. This simplifies the creation of Casanova.

There are no languages that use a strong typesystem and are fast enough for computing heavy applications, such as games.

\subsection{My contributions}
I have worked on MC with the main purpose of testing the language's expressivity.
\begin{lstlisting}
hallo=000
\end{lstlisting}


what did i do:
\begin{itemize}
	\item standard library (ga langs alle dingin hierin)
	\item basic monads
	\item started parser multiple times
\end{itemize}


fixing bugs
lift operator
understand the MC language
did a thing with F\# to learn more about functional programming


%\chapter{my task}
%my task was testing the expressivity of MC. Specifically ``can MC describe itself?''

\section{proper intro}

Right now Meta Casanova (MC) has no base of usage. and has not been testedin any pactical enviroment. This means we have no way of knowing how proctical MC actually is. We will set up basic test criteria to find out how good MC in a practical enviroment is. After which we explain MC in detail and put it to the test. Our findings will then lead to a conclusion as to what state MC is in.

\begin{enumerate}
	\item Firstly we explain why we want MC to be tested for it's goodness and usefulness.
	\item Then we will explain on what criteria we will judge MC.
	\item We will then dive into MC, explain the language and with it go ovet the criteria and see how it holds up.
	\item After which we will suggest improvements on MC.
	\item Lastly we will conclude how good and usefull MC really is at this stage and how to procede from here.
\end{enumerate}
