\chapter{The Problem}
  For a programming language to be usable and good it needs to be tested on  the following criteria:\cite{khedker1997makes,graham2004hackers}

\begin{multicols}{3}
\begin{enumerate}
  \item Read- \& Writability
  \item Simplicity
  \item Definiteness
  \item Orthogonality
  \item Expressiveness
  \item Implementability
  \item Efficiency
  \item Libraries
  \item Time
  \item Hackability
  \item Succintness
  \item Redesign
  \item External Factors
\end{enumerate}
\end{multicols}
Let's touch on each of these briefly, before we connect them to MC.

\begin{multicols}{2}
\paragraph{Writability and Readability}
of a programming language determines how good the connection between man and machine is. The programmer must be able to write programs he understands easily, even after months or years of not looking at it. This makes debugging much easier.

Documentation is makes up the basic part. But to really be able to make a program reabable, even after a prolonged absence, the language needs to be it's own documentation. This can be done via a manner of ways:
\begin{itemize}
  \item Keywords
  \item Abbreviations and concise notation
  \item Comments
  \item Layout or format of programs
  \item No overuse of notation
\end{itemize}
These points will be explained later on with examples off MC.\cite{}

\paragraph{Simplicity}
of a language is decided by it's features. These should be easy to learn and remember. To make this more provable we will make use of these subpoints:
\begin{itemize}
  \item Structure
  \item Number of features
  \item Multiple ways of specification
  \item Multiple ways of expressing
\end{itemize}
We will touch on these briefly with concern towards MC.\cite{} But the main concept through which simplicity is achieved, is through basing the language around a few simple well-defined concepts.\cite{}

\paragraph{Definiteness}
makes it clear for which purpose the language is designed and how it is to be used. When a language has a clear view and definition of itself during creation, it will also be clear to the user for what purposes to use tho language.

\paragraph{Orthogonality}


\paragraph{Expressiveness}

\paragraph{Implementability}

\paragraph{Efficiency}

\paragraph{Libraries}
are a necessity for every programming language. Without them the language would be useless and the programmer would have to build every feature from the ground up. This makes programming in the language take a long time.

\paragraph{Time}
is something everything needs to build momentum and a stable userbase, even programming languages can't excape this. For now we won't be able to really test this, as MC is still in it's development stages. For this reason we will leave this out.

\paragraph{Hackability}
is the ability to bend the language to one's will or to form the language to one's needs. In other words how much can you do with the language and how malleable and versatile is it. This also depends on the ability, of the programmer, to think outside the pre-defined box of the language. We will try to see how well MC supports this type of thinking.

\paragraph{Succinctness}
makes a language more abstract. When you have to say less to make the computer do precisly what you want, is something very powerfull. It also makes the programs shorter and clearer to read.

\paragraph{Redesign}
enables the evolution of a program written in the language. This makes it easier to go from a rough prototype to a fully featured program.

\paragraph{External factors}
play a big role in the adaptation of a language. Without some use for the language, it might aswell not exist. Having a platform for the language plays a major part in the way it will be adopted as a standard. This doesn't have to be a physical platform, like UNIX or Windows. It can be a virtual platform, as is an already existing major library. The only problem with the latter is, that the new language has to have something making it stand out from the readilly available languages who implement the library.

\paragraph{}
Do keep in mind that most of these criteria can be subjective to the user of the language and are not 100\% provable. We will try to discuss the most objective parts for provability's sake.

First I will explain MC in more detail, after which it can be put to the test. I will see how far it already has the above described criteria and in which areas it still needs some improvement.

SKETCH STUFF

MC in it's basic form is not

create a standard framework which makes the language usefull

The language in it's basic form is not very use

%% google for what makes a programming language usefull
\end{multicols}
